Pipit on the Post: Proving Pre- and Post-conditions of Reactive Systems

### 2024-01-15 2024pipit/section/70-Related.tex
conclusion

Verifying the time-triggered CAN driver is trickier, as the restrictions placed on the triggers array --- that triggers are sorted by time-mark, there must be an adequate time-gap between a trigger and its next-enabled, and a trigger's time-mark must be greater-than-or-equal-to its index --- naturally require quantifiers.
As described in \autoref{s:evaluation}, the Kind2 does include experimental array and quantifier support, but is limited to verifying small arrays up to 8 or 16 triggers.
Additionally, due to the limitations on top-level array definitions, compiling the program with Lustre~V6 would result in multiple copies of the entire triggers array on the stack.

@@

% Although Pipit is embedded in \fstar{}, which itself supports full dependent types and refinement types, we have opted to for a comparatively weakly-typed language.
% Our stream type \emph{stream $\tau$} denotes a (finite prefix of a) stream of elements of simple type $\tau$, but says nothing about how the stream evolves over time.

>>>>

Verifying the time-triggered CAN driver is trickier, as the restrictions placed on the triggers array --- that triggers are sorted by time-mark, there must be an adequate time-gap between a trigger and its next-enabled, and a trigger's time-mark must be greater-than-or-equal-to its index --- naturally require quantifiers.
As described in \autoref{s:evaluation}, the Kind2 does include experimental array and quantifier support, but is limited to verifying small arrays up to 8 or 16 triggers.
Additionally, due to the limitations on top-level array definitions, compiling the program with Lustre~V6 would result in the entire triggers array being copied to the stack each iteration.

@@

% Our stream type \emph{stream $\tau$} denotes a (finite prefix of a) stream of elements of simple type $\tau$, but says nothing about how the stream evolves over time.
% Although Pipit is embedded in \fstar{}, which itself supports full dependent types and refinement types, we have opted to for a comparatively weakly-typed language.

### 2024-01-15 2024pipit/section/80-Conclusion.tex
ok
extraction
conclusion

Our preliminary results show that \fstar{}'s proof automation and code extraction are suitable for verifying reactive systems and executing them in real-time; these results still require further work.
Next, we intend to verify the imperative code generation.
To verify large programs, we also need some way to separately prove smaller pieces which can then be composed together, such as contracts~\cite{champion2016kind2}.
Finally, we need to evaluate Pipit on larger control systems before extending the language to support more features, such as Lustre's clocks for describing partially-defined streams~\cite{caspi1995functional}.

% TODO
% Future work: verify imperative code generation; verify CSE; case studies: antilock braking; clocks \cite{caspi1995functional}; contracts \cite{champion2016kind2}.

>>>>

We have presented Pipit, a verified compiler and proof system for reactive systems.
Our implementation of the TTCAN driver logic shows that Pipit can express programs which are currently unsupported by other verified Lustre compilers.
Pipit can also verify high-level program properties which are difficult to express and prove in existing Lustre model-checkers.
Our development includes verified translations to both abstract and executable transition systems; both are shown to preserve the dynamic semantics.
We also introduced a checked semantics, which describes the proof obligations of a program; in future work, we intend to verify that any translated property the user proves on the transition system holds for the original property on the checked semantics.

In the future, we intend to increase the expressivity of Pipit by adding \emph{clocks}, which are used to describe partially-defined streams~\cite{caspi1995functional}.
Clocks are important for composing complex systems together and avoiding unnecessary computation.

We are interested in further pursuing the intersection of model-checking with interactive theorem proving.
A smart-contract called Djed \cite{zahnentferner2023djed} currently uses a mixture of Kind2 \cite{champion2016kind2} and manual Isabelle/HOL proofs to show that the contract is well-behaved.
In future work, we would like to further investigate whether Pipit's integration of streaming proofs with \fstar{}'s automated proof system would be able to provide similar proofs, without introducing any semantic gap between the two systems.
